A Model for Extracting Time Delays from Noisy Time Series Data

### 2013-08-25 doc/Paper1/ms.tex
Different notation for error bar boosts

\begin{table*}
\begin{tabular}{lll}
\hline
{\bf Parameter} &	{\bf Meaning}			& {\bf Prior}\\
\hline
$\tau_2$	&	Time delay of image 2	&	Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform$(m_{\rm min}, m_{\rm max})$\\
$\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
$L_q$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
$\{S_1, S_2\}$	&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
$\{\ell_1, \ell_2\}$ &	Length scale of image-specific variations & LogUniform$(\ell_{\rm min}, \ell_{\rm max})$\\
$\alpha$	&	Smoothness of image-specific variations	& Uniform$(1, 2)$\\
$f_{\rm bad}$	&	Fraction of ``bad'' data points	& Uniform$(0,1)$\\
$\kappa_1$	&	Error-bar boost for ``good'' data & $\frac{1}{2}\delta(\kappa_1 - 1) + \frac{1}{2}$LogUniform$(1,100)$\\
$\kappa_2$	&	Error-bar boost for ``bad'' data & LogUniform$(\kappa_1, 100\kappa_1)$\\
$\{b_i\}_{i=1}^N$ &	Badness hidden variables for each data point & Uniform$(0,1)$\\
\hline
{\bf Derived Parameter} & {\bf Meaning}	&	{\bf Definition}\\
\hline
$S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
$\{B_i\}_{i=1}^N$	&	Badness flag for each data point & $\mathds{1}\left(b_i \leq f_{\rm bad}\right)$\\
\hline
{\bf Data}	&	{\bf Meaning}			& {\bf Sampling Distribution}\\
\hline
$\{Y_1, ..., Y_N\}$ &	Noisy magnitude measurements	& $\mathcal{N}\left(\mathbf{m}(\params), \mathbf{C}(\params)\right)$\\
\hline
{\bf Prior Information} & {\bf Meaning}			&\\
\hline
$N$		&	Number of measurements (all images together)\\
$\{t_1, ..., t_N\}$ &	Measurement times\\
$\{k_1, ..., k_N\}$ &	Flags identifying which image was observed
\end{tabular}
\caption{A full list of parameters, prior information, and data in the
FLOTSAM model. If there are more than two images, extra parameters are
added accordingly.\label{tab:prior}}
\end{table*}

@@

In summary, the parameter vector is
\begin{equation}
\left\{ \{\tau\}, \{m\}, \sigma_{\rm QSO}, L_{\rm QSO}, \{\sigma_{\rm micro}\}
\{L_{\rm micro}\}, \alpha_{\rm micro}, \kappa \right\}
\end{equation}
where parameters in braces denote a set (one for each QSO image).  For a
double image lens, the total number of parameters in the model is~12. For a
quad, this number is~20. A set of uninformative prior
distributions for these parameters are listed in Table~\ref{priors}; we assign
these priors in the analysis described in the first part of this paper. Later,
we will show how a more informative set of priors can be derived, and used to
improve the precision of the time delay inferences {\bf (WE HOPE)}.

@@

\section{GP Covariance Function}
Recall that the noise-free light curves of the two images are given by:
\begin{eqnarray}
y_1(t) &=& m_1 + q(t - \tau_1) + \mu_1(t) \\
y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t)
\end{eqnarray}
where $\tau_1 = 0$.
Using this, and the model assumptions discussed in the paper, we can obtain the
covariance of image $i$ at time $t_1$ with image $j$ at time $t_2$, where
$i \neq j$:
\begin{eqnarray}
\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]
&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]\\
&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
\left(q(t_2 - \tau_j) + \mu_j(t_2)\right)\right]\\
&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_j) + q(t_1 - \tau_i)\mu_j(t_2)
+ \mu_i(t_1)q(t_2 - \tau_j) + \mu_i(t_1)\mu_j(t_2)
\right]\\
&=& S_q\exp
\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right].
\end{eqnarray}
We can also obtain the covariance of image $i$ measured at two times $t_1$ and
$t_2$:
\begin{eqnarray}
\textnormal{Cov}\left[y_i(t_1), y_i(t_2)\right]
&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_i(t_2) - m_i\right)\right]\\
&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
\left(q(t_2 - \tau_i) + \mu_i(t_2)\right)\right]\\
&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_i) + q(t_1 - \tau_i)\mu_i(t_2)
+ \mu_i(t_1)q(t_2 - \tau_i) + \mu_i(t_1)\mu_i(t_2)
\right]\\
&=& S_q\exp
\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right] +
S_i\exp\left[
-\left(\frac{|t_2 - t_1|
}{\ell_i}\right)^\alpha
\right]
.
\end{eqnarray}
The covariance between a data point $Y_i$ and another data point $Y_j$ is
given by the covariance of the underlying noise-free light curves at times $t_i$
and $t_j$, plus a noise term:
\begin{eqnarray}
\textnormal{Cov}\left(
Y_i, Y_j
\right) = \textnormal{Cov}\left(y_{k_i}(t_i), y_{k_j}(t_j)\right)
+ \delta_{ij}\left[\left[\kappa_1 + (\kappa_2 - \kappa_1)\mathds{1}(B_i = 1)\right]\sigma_i\right]^2
\end{eqnarray}
where $\sigma_i$ is the reported error bar given with the data. The coefficient
in front of $\sigma_i$ implies that the error bar should be multiplied by
$\kappa_1$ if this is a ``good'' data point, or by $\kappa_2$ if it is a ``bad''
data point.

>>>>

\begin{table*}
\begin{tabular}{lll}
\hline
{\bf Parameter} &	{\bf Meaning}			& {\bf Prior}\\
\hline
$\tau_2$	&	Time delay of image 2	&	Cauchy$\left(0, 0.1\tau_{\rm range}\right)T\left(\tau_{\rm min},\tau_{\rm max}\right)$\\
$\{m_1, m_2\}$	&	Mean magnitude for each image	& Uniform$(m_{\rm min}, m_{\rm max})$\\
$\beta$		&	Size of QSO variations		& LogUniform$(\beta_{\rm min}, \beta_{\rm max})$\\
$L_q$		&	Length scale of QSO variations	& LogUniform$(L_{\rm min}, L_{\rm max})$\\
$\{S_1, S_2\}$	&	Size of image-specific variations	& LogUniform$\left(A_{\rm min}, A_{\rm max}\right)$\\
$\{\ell_1, \ell_2\}$ &	Length scale of image-specific variations & LogUniform$(\ell_{\rm min}, \ell_{\rm max})$\\
$\alpha$	&	Smoothness of image-specific variations	& Uniform$(1, 2)$\\
$f_{\rm bad}$	&	Fraction of ``bad'' data points	& Uniform$(0,1)$\\
$\kappa_1$	&	Error-bar boost for ``good'' data & $\frac{1}{2}\delta(\kappa_1 - 1) + \frac{1}{2}$LogUniform$(1,100)$\\
$\kappa_2$	&	Error-bar boost for ``bad'' data & LogUniform$(\kappa_1, 100\kappa_1)$\\
$\{b_i\}_{i=1}^N$ &	Badness hidden variables for each data point & Uniform$(0,1)$\\
\hline
{\bf Derived Parameter} & {\bf Meaning}	&	{\bf Definition}\\
\hline
$S_{q}$	&	Standard deviation of QSO variability	& $\beta\sqrt{L}$\\
$\{a_i\}_{i=1}^N$	&	Errorbar multiplier for each data point & $\left\{\begin{array}{lr}\kappa_1, & b_i > f_{\rm bad}\\ \kappa_2, & b_i \leq f_{\rm bad}\end{array}\right.$\\
\hline
{\bf Data}	&	{\bf Meaning}			& {\bf Sampling Distribution}\\
\hline
$\{Y_1, ..., Y_N\}$ &	Noisy magnitude measurements	& $\mathcal{N}\left(\mathbf{m}(\params), \mathbf{C}(\params)\right)$\\
\hline
{\bf Prior Information} & {\bf Meaning}			&\\
\hline
$N$		&	Number of measurements (all images together)\\
$\{t_1, ..., t_N\}$ &	Measurement times\\
$\{k_1, ..., k_N\}$ &	Flags identifying which image was observed
\end{tabular}
\caption{A full list of parameters, prior information, and data in the
FLOTSAM model. If there are more than two images, extra parameters are
added accordingly.\label{tab:prior}}
\end{table*}

@@

@@

\section{GP Covariance Function}
Recall that the noise-free light curves of the two images are given by:
\begin{eqnarray}
y_1(t) &=& m_1 + q(t - \tau_1) + \mu_1(t) \\
y_2(t) &=& m_2 + q(t - \tau_2) + \mu_2(t)
\end{eqnarray}
where $\tau_1 = 0$.
Using this, and the model assumptions discussed in the paper, we can obtain the
covariance of image $i$ at time $t_1$ with image $j$ at time $t_2$, where
$i \neq j$. Here, $t_1$ and $t_2$ stand for any two times.
\begin{eqnarray}
\textnormal{Cov}\left[y_i(t_1), y_j(t_2)\right]
&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_j(t_2) - m_j\right)\right]\\
&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
\left(q(t_2 - \tau_j) + \mu_j(t_2)\right)\right]\\
&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_j) + q(t_1 - \tau_i)\mu_j(t_2)
+ \mu_i(t_1)q(t_2 - \tau_j) + \mu_i(t_1)\mu_j(t_2)
\right]\\
&=& S_q\exp
\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right].
\end{eqnarray}
We can also obtain the covariance of image $i$ measured at two times $t_1$ and
$t_2$:
\begin{eqnarray}
\textnormal{Cov}\left[y_i(t_1), y_i(t_2)\right]
&=& \mathds{E}\left[\left(y_i(t_1) - m_i\right)\left(y_i(t_2) - m_i\right)\right]\\
&=& \mathds{E}\left[\left(q(t_1 - \tau_i) + \mu_i(t_1)\right)
\left(q(t_2 - \tau_i) + \mu_i(t_2)\right)\right]\\
&=& \mathds{E}\left[q(t_1 - \tau_i)q(t_2 - \tau_i) + q(t_1 - \tau_i)\mu_i(t_2)
+ \mu_i(t_1)q(t_2 - \tau_i) + \mu_i(t_1)\mu_i(t_2)
\right]\\
&=& S_q\exp
\left[-\frac{|(t_2 -\tau_j)- (t_1-\tau_i)|}{L_q}\right] +
S_i\exp\left[
-\left(\frac{|t_2 - t_1|
}{\ell_i}\right)^\alpha
\right]
.
\end{eqnarray}
The covariance between a data point $Y_i$ and another data point $Y_j$ is
given by the covariance of the underlying noise-free light curves at times $t_i$
and $t_j$, plus a noise term:
\begin{eqnarray}
\textnormal{Cov}\left(
Y_i, Y_j
\right) = \textnormal{Cov}\left(y_{k_i}(t_i), y_{k_j}(t_j)\right)
+ \delta_{ij}\left(a_i\sigma_i\right)^2
\end{eqnarray}
where $\sigma_i$ is the reported error bar given with the data. The coefficient
$a_i$
in front of $\sigma_i$ is the robustness feature that depends on $\kappa_1$,
$\kappa_2$, and the badness variable for data point $i$.
